---
title: "ECOFF Finder Bayesian Approach"
author: "Tim Föckersperger, Katrin Leberfinger, Cornelia Volaucnik"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette_ECOFFBayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This documents is an introduction to the package `ECOFFBayes`. The aim of this package is to find cutoff points between resistant and non-resistant antibiotics. 

Therefore, first load the package `ECOFFBayes`.
```{r}
library(ECOFFBayes)
```

# Simulated Data

The package provides simulated data to try out the different functions.

```{r}
data("Antibiotics")
```

If you do not have any prior knowledge just apply this function to your data and it will come up with the number of components and the estimated parameter.

```{r, results = "hide"}
bayes.density <- TheBayesteApproach(Antibiotics)
```

If you have prior knowledge for your data, you can include this either through some prior parameter and/or through fixing the number of components k.

In this case, both information are included.

```{r, results = "hide"}
prior <- list(a.prior = 1, 
              b.prior = 1, 
              e0.prior = 1, 
              b0.prior = 5,
              B0.prior = 0.01, 
              c0.prior = 5, 
              C0.prior = 10)
k <- 3
bayes.density.with.prior <- TheBayesteApproach(data = Antibiotics,
                                               prior = prior, 
                                               k = k)
```

The output is of type `TheBayesteApproach` and includes the following components:

- number of components `k`
```{r}
bayes.density$k
bayes.density.with.prior$k
```

- parameter of the normal components
```{r}
bayes.density$normal.params
bayes.density.with.prior$normal.params
```

- parameter of binomial component
```{r}
bayes.density$binom.res
bayes.density.with.prior$binom.res
```

As we are also interested in the cutoff point between wildtype and non-wildtype bacteria (ECOFF), we can also calculate this value.
Therefore we 

```{r}
FindECOFF(bayes.density, 
          pi.level = 0.3, 
          quantile = 0.01)
FindECOFF(bayes.density.with.prior, 
          pi.level = 0.3, 
          quantile = 0.01)
```

For this object, also a plotting function is provided, which plots the histogram of the data and the estimated density.

A line that shows the ECOFF value can also be included in the graph. Therefore the parameter `ECOFF`, which is by default `FALSE`, needs to be set to `TRUE`.
Also here the `quantile` and `pi.level`parameter can be changed.

```{r, fig.height=5, fig.width=7, warning=FALSE}
plot(bayes.density,
     ECOFF = TRUE,
     quantile = 0.01,
     pi.level = 0.3)
```

```{r, fig.height=5, fig.width=7, warning=FALSE}
plot(bayes.density.with.prior,
     ECOFF = TRUE,
     quantile = 0.01,
     pi.level = 0.3)
```

# Data: EUCASTData

Get the data of the zone diameter from the external package `EUCASTData`. 

```{r}
data("ZD", package = "EUCASTData")
```

The code for three examples is included. Due to the required run-times for larger samples only the results for the second example are included, while the other examples are not evaluated.

## Example 1: 

For one example dataset, first only selected the antimicrobial and bacteria you are interested in.
Afterwards, you can run you analysis using the following functions:

```{r, eval = FALSE}
data1 <- subset(ZD, 
                Antimicrobial == "Ampicillin" & 
                  Bacterium == "Escherichia coli", 
                grepl("^Z", colnames(ZD)))
res1 <- TheBayesteApproach(data = data1)
FindECOFF(res1
          pi.level = 0.3, 
          quantile = 0.01)
plot(res1,
     ECOFF = TRUE,
     quantile = 0.01,
     pi.level = 0.3)
```

## Example 2: 

```{r, results = "hide"}
data2 <- subset(ZD, 
                Antimicrobial == "Piperacillin" & 
                  Bacterium == "Escherichia coli", 
                grepl("^Z", colnames(ZD)))
res2 <- TheBayesteApproach(data = data2)
```{r}
FindECOFF(res2, 
          pi.level = 0.3, 
          quantile = 0.01)
```{r, fig.height=5, fig.width=7, warning=FALSE}
plot(res2,
     ECOFF = TRUE,
     quantile = 0.01,
     pi.level = 0.3)
```

## Example 3: 

```{r, eval = FALSE}
data3 <- subset(ZD, 
                Antimicrobial == "Mecillinam" & 
                  Bacterium == "Escherichia coli", 
                grepl("^Z", colnames(ZD)))
res3 <- TheBayesteApproach(data = data3)
FindECOFF(res3, 
          pi.level = 0.3, 
          quantile = 0.01)
plot(res3,
     ECOFF = TRUE,
     quantile = 0.01,
     pi.level = 0.3)
```


