---
title: "ECOFF Finder Bayesian Approach"
author: "Tim Föckersperger, Katrin Leberfinger, Cornelia Volaucnik"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette_ECOFFBayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This documents is an introduction to the package `ECOFFBayes`. The aim of this package is to find cutoff points between resistant and non-resistant antibiotics. 

Therefore, first install and load the package ECOFFBayes.
```{r setup}
install("ECOFFBayes")
library(ECOFFBayes)
```

# Simulated Data

```{r data.sim}
load(Antibiotics)
```


```{r ex1.sim}
bayes.density <- TheBayesteApproach(Antibiotics)
```

The output is of type `TheBayesteApproach` and includes the following components:

- number of components `k`
```{r ex2.sim}
bayes.density$k
```

- parameter of the normal components
```{r ex3.sim}
bayes.density$normal.params
```

- parameter of binomial component
```{r ex4.sim}
bayes.density$biomial.params
```

For this object, also a plotting function is provided, which plots the histogram of the data and the estimated density.

```{r ex5.sim}
plot(bayes.density)
```

# Data: EUCASTData

Get the data of the zone diameter from the external package EUCASTData. 

```{r data, eval=FALSE}
library(ECOFFBayes)
data("ZD", package = "EUCASTData")
```

## Example 1: 

For one example dataset, first only selected the antimicrobial and bacteria you are interested in.

```{r example1, eval=FALSE}
data1 <- subset(ZD, 
                Antimicrobial == "Ampicillin" & 
                  Bacterium == "Escherichia coli", 
                grepl("^Z", colnames(ZD)))
```

If you do not have any prior knowledge just apply this function to your data and it will come up with the number of components and the estimated parameter.

```{r example1.1, eval=FALSE}
res <- TheBayesteApproach(data = data1)
```

The output is of type `TheBayesteApproach` and includes the following components:

- number of components `k`
```{r example1.2, eval=FALSE}
res1$k
```

- parameter of the normal components
```{r example1.3, eval=FALSE}
res1$normal.params
```

- parameter of binomial component
```{r example1.4, eval=FALSE}
res1$biomial.params
```

For this object, also a plotting function is provided, which plots the histogram of the data and the estimated density.

```{r example1.5, eval=FALSE}
plot(res2)
```

## Example 2: 

```{r example2, eval=FALSE}
data2 <- subset(ZD, 
                Antimicrobial == "Piperacillin" & 
                  Bacterium == "Escherichia coli", 
                grepl("^Z", colnames(ZD)))
```

If you have prior knowledge for your data, you can include this either through some prior parameter and/or through fixing the number of components k.

In this case, both information are included.
```{r example2.1, eval=FALSE}
prior <- list(a.prior = 1, b.prior = 1, e0.prior = 1, b0.prior = 5,
              B0.prior = 0.01, c0.prior = 5, C0.prior = 10)
k <- 2
res <- TheBayesteApproach(data = data2, prior = prior, k = k)
```


## Example 3: 

```{r example3, eval=FALSE}
data3 <- subset(ZD, 
                Antimicrobial == "Mecillinam" & 
                  Bacterium == "Escherichia coli", 
                grepl("^Z", colnames(ZD)))
```

```{r example3.1, eval=FALSE}
res3 <- TheBayesteApproach(data = data3)
```

As we are also interested in the cutoff point between wildtype and non-wildtype bacteria (ECOFF), we can also calculate this value.
Therefore we 

```{r example3.4, eval=FALSE}
FindECOFF(res3, 
          pi.level = 0.3, 
          quantile = 0.01)
```

A line that shows the ECOFF value can also be included in the graph. Therefore the parameter `ecoff`, which is by default `FALSE`, needs to be set to `TRUE`.
Also here the quantile can be changed.
```{r example3.5, eval=FALSE}
plot(res3, 
     ecoff = TRUE, 
     quantile = 0.01)
```

